Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 overall.tex
Address Jack's issue #14
Typo
Address Jack's issue #13

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
Titan contributed 3.9\% of the total of around 200K\mtnote{TODO: Fix this
number} cores available to ATLAS on Grid, Cloud, and HPC infrastructures
together. These figures confirms the relevance of supercomputers' resource
contribution to the LHC Run 2, especially when accounting for the amount of
unused backfill availability and the improvement of PanDA efficiency across the
experiment time window.

@@

Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
such, further hardware upgrades seem unlikely to improve significantly the
performance of PanDA Brokers. Nonetheless, the current load suggests that the
number of brokers per DTN could be increased. This would enable the submission
of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
to consume a higher percentage of backfill availability.

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events: the more events, the longer the simulation.
Not all events take the same time to be simulated:
% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
% of
1 event takes from $\approx$2 to $\approx$40 minutes to be simulated, with a
mean of $\approx$14 minutes. Considering that each worker node process up to 16
events concurrently, 100 events takes around 2 hours to process.

>>>>

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs.
% Comparatively, Titan contributed 3.9\% of the total of around
% 200K\mtnote{TODO: Fix this number} cores available to ATLAS on Grid, Cloud,
% and HPC infrastructures together.
These figures confirms the relevance of supercomputers' resource contribution to
the LHC Run 2, especially when accounting for the amount of unused backfill
availability and the improvement of PanDA efficiency across the experiment time
window.

@@

Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
such, further hardware upgrades seem unlikely to improve significantly the
performance of PanDA Brokers. Nonetheless, current DTN load suggests that the
number of brokers per DTN could be further increased. This would enable the
submission of a larger number of concurrent jobs to Titan's PBS queue, allowing
for PanDA to consume a higher percentage of backfill availability.

It should be noted that increasing the number of brokers requires also to
increase the number of concurrent jobs that can be submitted to Titan's queue.
Each broker submits 1 job at a time with at least 2 hours walltime. Backfill
availability is polled every 10 minutes, giving a theoretical maximum of 110
concurrent jobs submitted to Titan. Titan's policy was already changed in
September 2016, enabling the submission of 20 concurrent jobs and the deployment
of 20 PanDA brokers. Further study is required to understand the impact that a
larger amount of concurrent queued jobs would have on the overall efficiency of
Titan's scheduler.

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events: the more events, the longer the simulation.
Not all events take the same time to be simulated:
% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
% of
1 event takes from $\approx$2 to $\approx$40 minutes to be simulated, with a
mean of $\approx$14 minutes. Considering that each worker node process up to 16
events concurrently, 100 events takes around 2 hours to process. As such, PanDA brokers do not use backfill availability with less than 2 hours walltime.
